---
author: "Cynthia Kineza"
output: pdf_document
---

## The distribution of glucose value deviates from a normal distribution
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1234)
```
### Mixture models
```{r, echo=FALSE, warning=FALSE, message=FALSE}
glucose <- matrix(readRDS("~/Downloads/glucose.rds"))
colnames(glucose) <- "glucose_meas"
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
qqnorm(glucose)
qqline(glucose, col = 2,lwd=2,lty=1)
hist(glucose, main = "Distribution of data" )
```

## Implement a Gibbs sampler


Run at least 10,000 iterations. Compute and plot the autocorrelation of
$\theta_0$ and $\theta_1$ as well as the effective sample sizes.
```{r, echo=FALSE, warning=FALSE, message=FALSE}
#simulate zi
library(pscl)

# Here our iterations are called "s"

library(pscl)
a <- 1
b <- 1
vo <- 10
muo <- 120
sigmao <- sqrt(1000)
tauo <- sqrt(200)
y = NA

###initial value
p <- rbeta(1,a,b)
z <- rbinom(1, 1 , p)+1
theta1 <- rnorm(1, muo, tauo)
theta2 <- rnorm(1, muo, tauo)
variance_1 <- 1/rgamma(1, vo/2, vo*sigmao^2/2)
variance_2 <- 1/rgamma(1, vo/2, vo*sigmao^2/2)
min_theta <- min(theta1, theta2)
max_theta <- max(theta1, theta2)

glucose_final <- as.data.frame(rbind(c(theta1, theta2, variance_1, variance_2, p, y, min_theta, max_theta)))
colnames(glucose_final) <- c("theta1", "theta2", "var1", "var2", "p", "y", "theta1_tilde", "theta2_tilde")




if (z == 1){ 
  glucose_final[1,6] = rnorm(1, theta1, sqrt(variance_1))
} else {
  glucose_final[1,6] = rnorm(1, theta2, sqrt(variance_2))
}

zi <- vector(length=length(glucose))


set.seed(1)
for (s in 2:10000)
{
for (i in 1:length(glucose))
{

  y1 = dnorm(glucose[i],theta1, sd=sqrt(variance_1))
  y2 = dnorm(glucose[i], theta2, sd=sqrt(variance_2))
  zi[i] = rbinom(1,1,prob = (p*y1/(p*y1 + (1-p)*y2)))
}
  
  #summation of indicator function for zi = 1 = sum(2-zi)
   summation_1 = sum(zi==1)
   
   #summation of indicator function for zi = 2 = sum (zi-1)
   
   summation_2 = length(glucose) - summation_1
   
   #Lets now update the thetas from its full conditional
   
   theta1= rnorm(1,(sum(glucose*zi)*tauo^2 + variance_1*muo) / (variance_1 + tauo^2*summation_1), sd = sqrt(1/(summation_1/variance_1 + 1/tauo^2)))
   
   #generate the second theta here from its full conditional 

   theta2 = rnorm(1,(sum(glucose*(1-zi))*tauo^2 + variance_2*muo) /(variance_2 + tauo^2*summation_2), sd = sqrt(1/(summation_2/variance_2 + 1/tauo^2)))
    
   #generate sigm21 from its full conidtional, here i am using the first and second theta I obtained above
   
variance_1 <- 1/ rgamma(1, (vo+summation_1)/2, (sum(((glucose-theta1)^2)*zi) + sigmao^2*vo)/2)

    
   
      #generate sigm22 from its ful conditional, here i am using he first and second theta I obtained above
    variance_2 <- 1/ rgamma(1, (vo+summation_2)/2, (sum(((glucose-theta2)^2)*(1-zi)) + sigmao^2*vo)/2)
   
   
   #generate p (using the values I got for theta1_all,theta2_all,sigma21_end and sigma22_end)
   
   p = rbeta(1,a+summation_1,b+summation_2)
   
   
   #store this values in the pt's row in the matrix glucose_dat

   
   glucose_final[s,1] = theta1
   glucose_final[s,2] = theta2
   glucose_final[s,3] = variance_1
   glucose_final[s,4] = variance_2
   glucose_final[s,5] = p

   
    #updating z 
   
   z = rbinom(1,1,p) ### remember here I used 
   
   if (z==1)
   {
     glucose_final[s,6] = rnorm(1,theta1, sqrt(variance_1))
     
   }
   
   else
     
   {
     
     glucose_final[s,6] = rnorm(1,theta2, sqrt(variance_2))
   }
  
   glucose_final[s,7] = min(glucose_final[s,1],glucose_final[s,2]) #theta_min
   glucose_final[s,8] = max(glucose_final[s,1],glucose_final[s,2]) #theta_max
   

   
   
  
}
library(coda)
acf(glucose_final[,7], type = "correlation", plot=FALSE)
acf(glucose_final[,8], type = "correlation", plot=FALSE)
acf(glucose_final[,7], type = "correlation", main = "Autocorrelation for first theta")
acf(glucose_final[,8], type = "correlation", main = "Autocorrelation for second theta")

effectiveSize(glucose_final[,7])
effectiveSize(glucose_final[,8]) 

```

For each iterations `s` of the Gibbs sampler, sample


```{r, echo=FALSE, warning=FALSE, message=FALSE}
hist(glucose_final$y, main = "Histogram of sampling data y")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
df <- data.frame(parameter=as.numeric(glucose_final$theta1_tilde),
                 name=rep(c("theta1_min", "precision"), each=nrow(glucose_final)),
                 iter=rep(seq_len(nrow(glucose_final)),2))

ggplot(df, aes(iter, parameter)) +
  geom_line() +
  facet_wrap(~name, ncol=1, scales="free_y") +
  xlab("MCMC iteration") + ylab("")


df.1 <- data.frame(parameter=as.numeric(glucose_final$theta2_tilde),
                 name=rep(c("theta2_max", "precision"), each=nrow(glucose_final)),
                 iter=rep(seq_len(nrow(glucose_final)),2))

ggplot(df.1, aes(iter, parameter)) +
  geom_line() +
  facet_wrap(~name, ncol=1, scales="free_y") +
  xlab("MCMC iteration") + ylab("")


```

Joint posterior is difficult to sample directly, easier to sample from the full conditional distribution of each parameter. We used a Gibbs Sampler generates a dependent sequence of parameter values whose distribution converges to the target joint posterior distribution


The first thetas were arbitrarily chosen initial values, then the thetas generated in the gibbs sampler depended on the initial thetas. For any initial values, the chain is converging to the posterior.
Using 10,000 iterations, we used enough samples to approximate our posterior

The original glucose data does not follow a normal distribution (it has two modes, 100, 150), while the empirical distribution of $Y$ approximately follows normal distribution with 121 (approximately) as mean.
Based on the iteration results, the average of $\theta_1$ is 104.10  while the average of $\theta_2$ is 148.7, which is coherent to the two modes location of the histogram of glucose data.

